%%  pref-dev.tex,  version 06/03/19

%%%%%%%%%%%%%%%%%%
\section*{Preface}

\addcontentsline{toc}{part}{Preface}

The aim of these notes is to collect together, using a common notation
and right actions, information about crossed modules; cat1-groups;
crossed squares; cat2-groups; and related structures.

Inspiration has come from Ronnie Brown; Tim Porter; postgraduate students;
and the many visitors to the department in Bangor.

Help in the preparation of these notes has initially been 
provided by Murat Alp and Robert Rodrigues.
A seminar course for Gareth Evans and Richard Lewis during 2002/2003 
resulted in many corrections and additions.
Many additions were made when Murat Alp visited Bangor in July and August 2003.

After that, some material was added up until 2008. 
The notes were returned to in April 2017 when work started with Ronnie Brown 
of induced crossed squares and with Alper Odabas on \GAP\ code for cat2-groups. 
The notes were deposited on the \arXiv\ in February 2018. 

In parallel with these notes the \GAP\ packages \XMod\ and \Groupoids\ 
have been developed to perform computations with these structures. 
These packages may be obtained from the main \GAP\ package distribution at
\url{https://www.gap-system.org/Packages/} 
or from the \GitHub\ repositories at:
\url{https://gap-packages.github.io/groupoids/} and
\url{https://gap-packages.github.io/xmod/}. 

If you have a question relating to either of these packages, encounter any problems, 
or have a suggestion for extending the packages in any way, please 
make an issue at: \url{https://github.com/gap-packages/groupoids/issues/new} 
or \url{https://github.com/gap-packages/xmod/issues/new}. 

\bigskip
The main sources used are (up until 2008): 
Alp \& Wensley \cite{alp:wens-ijac} (2000); 
Brown \& Gilbert \cite{brow:gilb} (1989); 
Brown \& Huebschmann \cite{brow:hueb} (1982); 
Brown and \.{I}\c{c}en \cite{brow:icen} (2003); 
Brown \& Loday \cite{brow:lod} (1987); 
Ellis \cite{ellis-thesis} (1984); 
Ellis \& Steiner \cite{ell:st} (1987); 
Kamps \& Porter \cite{kamps:port} (2002); 
Norrie \cite{norrie-thesis} (1987) and 
Porter \cite{porter:top} (1993).

\bigskip

This is the development version of the notes, containing lots of
comments and suggestions of items which need to be corrected, added
or improved.

Much of the later sections were to form part of a paper 
on the actor of a crossed square, but this was never completed. 
