%%  pref-dev.tex,  version 22/04/17

%%%%%%%%%%%%%%%%%%
\section*{Preface}

\addcontentsline{toc}{part}{Preface}

The aim of these notes is to collect together, using a common notation
and right actions, information about crossed modules; cat1-groups;
crossed squares; cat2-groups; and related structures.

Inspiration has come from Ronnie Brown; Tim Porter; postgraduate students;
and the many visitors to the department in Bangor.

Help in the preparation of these notes has initially been 
provided by Murat Alp and Robert Rodrigues.
A seminar course for Gareth Evans and Richard Lewis during 2002/2003 
resulted in many corrections and additions.
Many additions were made when Murat Alp visited Bangor in July and August 2003.

After that, material was added up until 2008, 
and then a long gap when nothing happened. 
The notes were returned to in April 2017 when work started with Ronnie Brown 
of induced crossed squares and with Alper Odabas on GAP code for cat2-groups. 

\bigskip
The main sources used are (up until 2008):
\begin{itemize}
\item
M.~Alp and C.D.~Wensley,
\emph{Enumeration of cat1-groups of low order},
Int J Alg Comp 10 (2000) 407--424.

\item
R.~Brown and N.D.~Gilbert,
\emph{Algebraic models of 3-types and automorphism
structures for crossed modules},
Proc. London Math. Soc. (3) 59 (1989) 51--73.
\item
R.~Brown and J.~Huebschmann, \emph{Identities among relations},
in Low-dimensional topology, London Math. Soc. Lecture Note Series 48,
ed. R Brown and T L Thickstun, Cambridge University Press, 1982, pp.153--202.

\item
R.~Brown and  \.{I}.~\.{I}\c{c}en,
\emph{Homotopies and automorphisms of crossed modules of groupoids},
Applied Categorical Structures 11 (2003) 185--206.

\item
R.~Brown and  J.-L.~Loday,
\emph{Van Kampen theorems for diagram of spaces},
Topology 26 (1987) 311--335.

\item
G.J.~Ellis, 
\emph{Crossed modules and their higher dimensional analogues},
Ph.D. Thesis, University of Wales, Bangor, 1984.

\item
G.J.~Ellis and R.~Steiner,
\emph{Higher-dimensional crossed modules and the
homotopy groups of $(n+1)$-ads},
J Pure and Appl. Algebra 46 (1987) 117--136.

\item
K.H.~Kamps and T.~Porter,
\emph{2-groupoid enrichments in homotopy theory and algebra},
K-Theory 25 (2002) 373--409.

\item
K.J.~Norrie, 
\emph{Crossed modules and analogues of group theorems},
Ph.D. Thesis, King's College, London, January 1987.

\item
T. Porter,
\emph{$n$-types of simplicial groups and crossed $n$-cubes},
Topology 32 (1993) 5--24.

\item
(more to be added)
\end{itemize}

\bigskip

This is the development version of the notes, containing lots of
comments and suggestions of items which need to be corrected, added
or improved.

Much of the later sections were to form part of a paper 
on the actor of a crossed square, but this never happened. 
