%%  preface.tex,  version 12/06/09

%%%%%%%%%%%%%%%%%%
\section*{Preface}

\addcontentsline{toc}{part}{Preface}

The aim of these notes is to collect together, using a common notation
and right actions, information about crossed modules; cat1-groups;
crossed squares; and related structures.

Inspiration has come from Ronnie Brown; Tim Porter; postgraduate students;
and the many visitors to the department in Bangor.

Help in the preparation of these notes has initially been 
provided by Murat Alp and Robert Rodrigues.
A seminar course for Gareth Evans and Richard Lewis during 2002/2003
resulted in many corrections and additions.

{\bf This is work in progress}, so there are many comments in the text 
of the form\\ 
``{\bf [Such-and-such needs doing.]}''.

The \emph{hyperref} package is used from June 2009 to enable readers 
to skip around these notes. 

\bigskip
The main sources used are:
\begin{itemize}
\item
M.~Alp and C.D.~Wensley,
\emph{Enumeration of cat1-groups of low order},
Int J Alg Comp 10 (2000) 407--424.

\item
M.~Alp and C.D.~Wensley,
\emph{Automorphisms and homotopies of groupoids and crossed modules},
Bangor Preprint 08.05, Appl. Categor. Struct. (to appear).

\item
R.~Brown and N.D.~Gilbert,
\emph{Algebraic models of 3-types and automorphism
structures for crossed modules},
Proc. London Math. Soc. (3) 59 (1989) 51--73.
\item
R.~Brown and J.~Huebschmann, \emph{Identities among relations},
in Low-dimensional topology, London Math. Soc. Lecture Note Series 48,
ed. R Brown and T L Thickstun, Cambridge University Press, 1982, pp.153--202.

\item
R.~Brown and  \.{I}.~\.{I}\c{c}en,
\emph{Homotopies and automorphisms of crossed modules of groupoids},
Applied Categorical Structures 11 (2003) 185--206.

\item
R.~Brown and  J.-L.~Loday,
\emph{Van Kampen theorems for diagram of spaces},
Topology 26 (1987) 311--335.

\item
G.J.~Ellis, 
\emph{Crossed modules and their higher dimensional analogues},
Ph.D. Thesis, University of Wales, Bangor, 1984.

\item
G.J.~Ellis and R.~Steiner,
\emph{Higher-dimensional crossed modules and the
homotopy groups of $(n+1)$-ads},
J Pure and Appl. Algebra 46 (1987) 117--136.

\item
K.H.~Kamps and T.~Porter,
\emph{2-groupoid enrichments in homotopy theory and algebra},
K-Theory 25 (2002) 373--409.

\item
K.J.~Norrie, 
\emph{Crossed modules and analogues of group theorems},
Ph.D. Thesis, King's College, London, January 1987.

\item
T. Porter,
\emph{$n$-types of simplicial groups and crossed $n$-cubes},
Topology 32 (1993) 5--24.

\item
(more to be added)
\end{itemize}
